---
title: "Annova Analysis"

----


install.packages("ISwR")

```{r echo=TRUE}
library(ISwR)
attach(red.cell.folate)
summary(red.cell.folate)
```

## One-way analysis of variance
Variation between groups is labelled by the name of the grouping factor (ventilation), and variation within groups is labelled Residual. 

```{r echo=TRUE}
anova(lm(folate~ventilation)) 
```

# Two-way analysis of variance
To perform a two-way ANOVA, it is necessary to have data in one vector,
with the two classifying factors parallel to it. We consider an example
concerning heart rate after administration of enalaprilate.

```{r echo=TRUE}
attach(heart.rate)
heart.rate
```
If you look inside the heart.rate.R file in the data directory of the
ISwR package, you will see that the actual definition of the data frame is

```{r echo=TRUE}
heart.rate <- data.frame(hr = c(96,110,89,95,128,100,72,79,100,
92,106,86,78,124,98,68,75,106,
86,108,85,78,118,100,67,74,104,
92,114,83,83,118,94,71,74,102),
subj=gl(9,1,36),
time=gl(4,9,36,labels=c(0,30,60,120)))
```
The gl (generate levels) function is specially designed for generating patterned factors for balanced experimental designs. It has three arguments:
the number of levels, the block length (how many times each level should
repeat), and the total length of the result. The two patterns in the data
frame are thus

```{r echo=TRUE}
gl(9,1,36)
```
```{r echo=TRUE}
gl(4,9,36,labels=c(0,30,60,120))
```
Once the variables have been defined, the two-way analysis of variance is
specified simply by
```{r echo=TRUE}
anova(lm(hr~subj+time))
```
