---
title: "hypotesis testing"

---

## mtcars
```{r echo=TRUE}
summary(mtcars)
head(mtcars)
```

## linear regression
```{r echo=TRUE}
lm(mtcars$mpg~mtcars$cyl)
```
gives us the linear equation: mpg = 37.885 -2.876 cyl

## linear regression
```{r echo=TRUE}
summary(lm(mtcars$mpg~mtcars$cyl))
```

## Correlation Matrix
```{r echo=TRUE}
cor(mtcars[,1:6])
```

## Spearman Correlation
```{r echo=TRUE}
cor(mtcars$mpg, mtcars$cyl, method="spearman")
```

## Pearson Correlation
```{r echo=TRUE}
cor(mtcars$mpg, mtcars$cyl, method="pearson")
```

## Significance of Correlation Coefficients for Spearman
```{r echo=TRUE}
cor.test(mtcars$mpg, mtcars$cyl, method="spearman")
```

## Significance of Correlation Coefficients for Pearson
```{r echo=TRUE}
cor.test(mtcars$mpg, mtcars$cyl, method ="pearson")
```

## Visualizing Correlation
```{r echo=TRUE}
library("corrplot", lib.loc="~/R/win-library/3.2")
data(mtcars)
corrplot(cor(mtcars), method="ellipse")
```

## Wilcoxon Rank Sum Test
```{r echo=TRUE}
x<-c(235,225,190,188)
y<-c(180,169,180,185,178,182)
wilcox.test(x, y)
```

## Kruskal-Wallis Rank Sum Test
```{r echo=TRUE}
a<-c(65,87,73,79,81,69)
b<-c(75,69,83,81,72,79,90)
c<-c(59,78,67,62,83,76)
d<-c(94,89,80,88)
x <- c(a, b, c, d)
g <- factor(rep(1:4, c(6,7, 6, 4)), labels = c("1","2","3","4"))
kruskal.test(x,g)
```

##Friedman Rank Sum Test
```{r echo=TRUE}
ReactionTimes <- matrix(c(0.6, 0.7, 0.9, 0.5, 0.9, 1.1,
1.3, 0.7, 0.8, 0.7, 1.0, 0.8), nrow = 4, byrow = FALSE)
friedman.test(ReactionTimes)
```


