---
title: "Regression modelling( Multivariate, step-wise, polynomial Regression)"

---

Preparation:
```{r}
library(RCurl)
library(MASS) # stepwise regression
library(leaps) # all subsets regression
library(glmnet) # for regularized linear regression
```

## Linear Regression: Fitting the Model
```{r echo=TRUE}
plot(airquality$Wind, airquality$Temp)
```

## Linear Regression: Fitting the Model
```{r echo=TRUE}
model_ulm <- lm(Wind~Temp, data=airquality) 
model_ulm
```
## Linear Regression: Prediction
```{r echo=TRUE, eval=TRUE}
rn_train <- sample(nrow(airquality), 
                   floor(nrow(airquality)*0.8))
train <- airquality[rn_train,c("Wind","Temp")]
test <- airquality[-rn_train,]
model_ulm <- lm(Wind~Temp, data=train) 
prediction <- predict(model_ulm, interval="prediction", 
                      newdata =test)
```

## Linear Regression: Error Distribution
```{r echo=TRUE, fig.width=10, fig.height=10}
errors <- prediction[,"fit"] - test$Wind
hist(errors)
```

## Linear Regression: RMSE
```{r echo=TRUE}
sqrt(sum((prediction[,"fit"] - test$Wind)^2)/nrow(test))
```

## Linear Regression: PRED(25)
Find the percentage of cases with less than 25 percent error:
```{r echo=TRUE}
rel_change <- 1 - ((test$Wind - abs(errors)) / test$Wind)
table(rel_change<0.25)["TRUE"] / nrow(test)
```

## Multivariate Linear Regression
```{r}
model_mlr <- lm(Temp~Month+Wind, data=airquality) 
model_mlr
```

## Multivariate Linear Regression: Test Example
```{r echo=TRUE, eval=TRUE}
rn_train <- sample(nrow(airquality), 
                   floor(nrow(airquality)*0.8))
train <- airquality[rn_train,]
test <- airquality[-rn_train,]
model_mlr <- lm(Temp~Month+Wind, data=train) 
prediction <- predict(model_mlr, interval="prediction", 
                      newdata=test)
```

## Multivariate Linear Regression: RMSE

```{r echo=TRUE}
sqrt(sum((prediction[,"fit"] - test$Temp)^2)/nrow(test))
```

## Multivariate Linear Regression: PRED(10)
Find the percentage of cases with less than 10 percent error:
```{r echo=TRUE}
errors <- prediction[,"fit"] - test$Temp
rel_change <- 1 - ((test$Temp - abs(errors)) / test$Temp)
table(rel_change<0.10)["TRUE"] / nrow(test)
```

## Stepwise Linear Regression - Forward

Start from a null formula. Go forward.

```{r echo=TRUE}
full <- lm(mpg~wt+hp+disp+gear,data=mtcars)
null <- lm(mpg~1,data=mtcars)
stepF <- stepAIC(null, scope=list(lower=null, upper=full), 
                direction= "forward", trace=FALSE)
summary(stepF)
```

## Stepwise Linear Regression - Forward

Start from a null formula. Go forward. We set trace=TRUE to see all the steps.
```{r echo=TRUE}
full <- lm(mpg~wt+hp+disp+gear,data=mtcars)
null <- lm(mpg~1,data=mtcars)
stepF <- stepAIC(null, scope=list(lower=null, upper=full), 
                direction= "forward", trace=TRUE)
summary(stepF)
```

## Stepwise Linear Regression - Backward

Start from a full formula. Go back.

```{r echo=TRUE}
full <- lm(mpg~wt+hp+disp+gear,data=mtcars)
stepB <- stepAIC(full, direction= "backward", trace=FALSE)
summary(stepB)
```

## Stepwise Linear Regression - Backward

Start from a full formula. Go back. We set trace=TRUE to see all the steps.

```{r echo=TRUE}
full <- lm(mpg~wt+hp+disp+gear,data=mtcars)
stepB <- stepAIC(full, direction= "backward", trace=TRUE)
summary(stepB)
```

## Best Subset
Select the best subset of the variables through exhaustive search.

```{r}
subsets<-regsubsets(mpg~wt+hp+disp+gear,data=mtcars,
                    nbest=2)
sub.sum <- summary(subsets)
as.data.frame(sub.sum$outmat)
```

## Polynomial Regression
```{r}
library(ISwR)
attach(cystfibr)
lm(pemax~height+I(height^2))
pred.frame <- data.frame(height=seq(110,180,2))
lm.pemax.hq <- lm(pemax~height+I(height^2))
predict(lm.pemax.hq,interval="pred",newdata=pred.frame)
```

## Polynomial Regression
```{r}
pp <- predict(lm.pemax.hq,newdata=pred.frame,interval="pred")
pc <- predict(lm.pemax.hq,newdata=pred.frame,interval="conf")
plot(height,pemax,ylim=c(0,200))
matlines(pred.frame$height,pp,lty=c(1,2,2),col="black")
matlines(pred.frame$height,pc,lty=c(1,3,3),col="black")
```

